---
title: "Causes of death"
author:
- familyname: Cheung
  othernames: Yan Chui Lucia
  address: Monash Universtidy
  email: yche0164@student.monash.edu
  correspondingauthor: true
  qualifications:  Master of Business Analytics
- familyname: Soh
  othernames: Min Min
  address: Monash Universtidy
  email : mmsoh1@student.monash.edu
  qualifications: Master of Business Analytics
- familyname: Meer
  othernames: Sanna
  address: Monash University
  email: smee0005@student.monash.edu
  correspondingauthor: true
  qualifications:  Master of Business Analytics
- familyname: Imuan
  othernames: Aphiaut
  address: Monash Universtidy
  email: aimu0001@student.monash.edu
  correspondingauthor: true
  qualifications:  Master of Business Analytics
department: ETC5513 Group 4
organization: Monash University
bibliography: References.bib
biblio-style: authoryear-comp
linestretch: 1.5
output:
  bookdown::pdf_document2:
    template: monashreport2.tex
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    includes:
      in_header: preamble.tex
    keep_tex: yes
    number_sections: yes
    citation_package: biblatex
    toc: false
---

```{r , echo = FALSE, message = FALSE, warning = FALSE}
# Libraries
library(tidyverse)
library(readr)
library(kableExtra)
library(bookdown)
library(gridExtra)
library(tinytex)
library(knitr)
```


# Introduction 



# Library used

The library packages used in our research consist of: 

- tidyverse (@tidyverse)
- readr (@readr)
- kableExtra  (@kableExtra)
- Bookdown (@bookdown @bookdown1)
- gridExtra (@gridExtra)
- Tinytex (@tinytex @tinytex1)
- Knitr (@knitr @knitr1)

```{r readdata, , echo = FALSE, message = FALSE, warning = FALSE}
Data <- read_csv("Data/annual-number-of-deaths-by-cause.csv")

```

\clearpage

\section*{}

## YanChuiLucia_Cheung


\section*{}
## MinMin_Soh


\clearpage





\section*{}

## Sanna_Meer



\section*{Deaths - Cardiovascular diseases}

## Aphiaut_Imuan  

The research question for analysis in this section are :  
- How is **the trend of Cardiovascular diseases** changed across the years in the **world**? 
- What are **the 5 countries with the highest death** from Cardiovascular diseases?
- What are the healthy lifestyle factors for Cardiovascular diseases?  


```{r, echo=FALSE, message=FALSE}
cardio <- Data %>%
  rename("cardio_num" = "Deaths - Cardiovascular diseases - Sex: Both - Age: All Ages (Number)") %>%
  rename("country" = Entity)%>%
  select(country,Year, cardio_num) %>%
  filter(!country %in% c("G20", "World Bank Upper Middle Income",
                         "East Asia & Pacific (WB)", "World Bank Lower Middle Income",
                         "Western Pacific Region (WHO)", "European Region (WHO)",
                         "Europe & Central Asia (WB)", "South-East Asia Region (WHO)",
                         "OECD Countries", "World Bank High Income", "South Asia (WB)",
                         "Region of the Americas (WHO)", "Eastern Mediterranean Region (WHO)",
                         "Latin America & Caribbean (WB)", "African Region (WHO)",
                         "Sub-Saharan Africa (WB)", "North America (WB)",
                         "Middle East & North Africa (WB)", "World Bank Low Incom", "World Bank Low Income"))

```

```{r worldtrendcardio, echo=FALSE, echo=FALSE ,out.width = "80%", fig.height = 5, fig.align='center', fig.cap= "The trend of Cardiovascular diseases in the world", warning=FALSE, fig.pos = 'H', message=FALSE}
cardio_world <- cardio %>%
  filter(country == "World")

graph_world <-  cardio_world %>%
  group_by(country)%>%
  ggplot() +
  geom_line(aes(x = Year, y = cardio_num, color = country)) +
  ylab("Number of death") + xlab("Year") + ggtitle("Cardiovascular diseases world trend")+ scale_y_continuous(labels = scales::comma)

graph_world
```

```{r tabletop5cardio, echo=FALSE, echo=FALSE ,out.width = "80%", fig.height = 5, fig.align='center', warning=FALSE, fig.pos = 'H', message=FALSE}
cardio_top5 <- cardio %>%
  filter(country %in% c( "China", "India", "Russia", "United States", "Indonesia"))


table_cardio <-  cardio_top5 %>%
  group_by(country)%>%
  summarise(mean = mean(cardio_num, na.rm = TRUE),
            max = max(cardio_num, na.rm = TRUE)) %>%
  arrange(-max)

table_cardio%>%
    kable(caption = "The top 5 countries by death of Cardiovascular diseases")%>%
    kable_styling(full_width = FALSE, position = "center",latex_options = "hold_position")

```
```{r graphtop5cardio, echo=FALSE ,out.width = "80%", fig.height = 5, fig.align='center', fig.cap= "The 5 countries with the highest death from Cardiovascular diseases", warning=FALSE, fig.pos = 'H', message=FALSE}
graph <-  cardio_top5 %>%
  group_by(country)%>%
  ggplot() +
  geom_line(aes(x = Year, y = cardio_num, color = country)) +
  ylab("Number of death") + xlab("Year") + ggtitle("The 5 countries with the highest death from Cardiovascular diseases") +scale_y_continuous(labels = scales::comma)


graph
```

The number of deaths from cardiovascular diseases from 1990 to 2019 in Figure \@ref(fig:worldtrendcardio) shows a significant positive trend and is likely to continue to increase. Table \@ref(tab:tabletop5cardio) illustrates the top 5 countries with the highest death from cardiovascular diseases including China, India, Russia, the United States, and Indonesia, respectively.  Moreover, @zhang_2021_healthy investigated people dying from cardiovascular disease in China are described by the high-salt food, high cholesterol, and high smoking rate. @yang_2015_cardiovascular argue that around 17% of the number of death by cardiovascular diseases were due to ambient temperature and the number of death from cold temperatures more than heat temperatures. These studies can support the result of China in Figure \@ref(fig:graphtop5cardio).
Furthermore, figure 2 shows India and Indonesia are positive trend like China because 35 % of Indian adults are smoking (@prabhakaran_2016_cardiovascular) and  26.6 % of Indonesian adults aged 40 years (@maharani_2019_cardiovascular).

In the contrast, Russia has decreasing trend since 2005 because Russia has National Priority Project ‘Health’ to increase the modern treatment and prevention of cardiovascular diseases (@pogosova_2016_positive). Moreover, the proportions of smokers in Russia are less than 50 percent which makes the decreasing trend in the number of deaths in Russia (@lakunchykova_2020_why). Furthermore, @zhang_2021_healthy investigate healthy lifestyle factors such as non-smoking, exercise, eating some healthy food, and don’t drink alcohol can reduce the mortality rate of cardiovascular in the USA that is related with our result that the trend of death starts to decrease around 2004. However, it will back to increase because some US populations have been unable access to cardiovascular treatment(@mensah_2017_decline). 



\clearpage

# Conclusion :

Cardiovascular part

The government should create a project or have some organization to support and prevent cardiovascular diseases. Moreover, they should make a policy for people to reduce or stop smoking such as increasing the tax on cigarettes for making people can't afford cigarettes.

\clearpage