---
title: "Causes of death"
author:
- familyname: Cheung
  othernames: Yan Chui Lucia
  address: Monash Universtidy
  email: yche0164@student.monash.edu
  correspondingauthor: true
  qualifications:  Master of Business Analytics
- familyname: Soh
  othernames: Min Min
  address: Monash Universtidy
  email : mmsoh1@student.monash.edu
  qualifications: Master of Business Analytics
- familyname: Meer
  othernames: Sanna
  address: Monash University
  email: smee0005@student.monash.edu
  correspondingauthor: true
  qualifications:  Master of Business Analytics
- familyname: Imuan
  othernames: Aphiaut
  address: Monash Universtidy
  email: aimu0001@student.monash.edu
  correspondingauthor: true
  qualifications:  Master of Business Analytics
department: ETC5513 Group 4
organization: Monash University
bibliography: References.bib
biblio-style: authoryear-comp
linestretch: 1.5
output:
  bookdown::pdf_document2:
    template: monashreport2.tex
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    includes:
      in_header: preamble.tex
    keep_tex: yes
    number_sections: yes
    citation_package: biblatex
    toc: false
---

```{r , echo = FALSE, message = FALSE, warning = FALSE}
# Libraries
library(tidyverse)
library(readr)
library(kableExtra)
library(bookdown)
library(gridExtra)
library(tinytex)
library(knitr)
```


# Introduction 



# Library used

The library packages used in our research consist of: 

- tidyverse (@tidyverse)
- readr (@readr)
- kableExtra  (@kableExtra)
- Bookdown (@bookdown @bookdown1)
- gridExtra (@gridExtra)
- Tinytex (@tinytex @tinytex1)
- Knitr (@knitr @knitr1)

```{r readdata, , echo = FALSE, message = FALSE, warning = FALSE}
Data <- read_csv("Data/annual-number-of-deaths-by-cause.csv")

```

\clearpage

\section*{}

## YanChuiLucia_Cheung


## MinMin_Soh

# Deaths - HIV/AIDS

Research Questions: 

- How is the death rates of **HIV / AIDS** change in the world ? 
- What is the top 5 countries with the highest death rates from **HIV / AIDS** and the death trends for those countries?
- What are the preventive measures for **HIV / AIDS** ? 

\clearpage

## Death rates of **HIV / AIDS** in the world  

```{r HIV_AIDS,  echo = FALSE, message = FALSE, warning = FALSE }
HIV_AIDS <- Data %>% 
  select( Entity, Code, Year, `Deaths - HIV/AIDS - Sex: Both - Age: All Ages (Number)`) %>%
  rename(Deaths_HIV_AIDS = `Deaths - HIV/AIDS - Sex: Both - Age: All Ages (Number)` )%>%
  filter(!is.na(Deaths_HIV_AIDS))
                
                 
```



```{r WorldHIVtrend, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Number of Deaths from HIV/AIDS in the World" }

HIV_AIDS %>% filter (Entity == "World")%>%
  ggplot(aes(x= Year, 
             y = Deaths_HIV_AIDS))+
  geom_line(linetype = "dashed", color = "blue", size =1 ) +
  geom_point(color = "blue", size = 3)+
  labs(y = "Number of Deaths from HIV/AIDS",
       x = "Year")+
  ggtitle("Deaths rates from HIV/AIDS in the world")+
  scale_y_continuous(labels = scales::comma)+
  theme_minimal()

```

Figure \@ref(fig:WorldHIVtrend)  depicts the global death rate of HIV/AIDS over time. As observed, the death rate followed an increasing trend from 1990 to around 2005, and then decreased from 2005 onwards. 

\clearpage

## The 5 countries with the highest death rates from **HIV / AIDS**

```{r  average_death_HIV,  echo = FALSE, message = FALSE, warning = FALSE}
#disable scientific notation
options(scipen=999)

#calculate average death rates per nation
HIV_AIDS_nation <- HIV_AIDS %>% 
  filter(!Entity %in% c("Asia", "Africa", "African Region", "Central Asia", "Central Europe", "Central Europe, Eastern Europe, and Central Asia",
                        "Central Latin America", "Central sub-Saharan Africa", "Commonwealth High Income", "Commonwealth Low Income", 
                        "Commonwealth Middle Income" , "East Asia","East Asia & Pacific - World Bank region", "Eastern Europe", 
                        "Eastern Mediterranean Region", "Eastern sub-Saharan Africa" ,"Europe & Central Asia - World Bank region", 
                        "European Region", "European Union" , "G20", "Low SDI", "High SDI", "Low-middle SDI",  "High-income", 
                        "High-income Asia Pacific", "High-income North America" , "High-middle SDI","Latin America & Caribbean - World Bank region" ,
                        "Middle SDI","Commonwealth", "African Union", "Southern sub-Saharan Africa", "Region of the Americas",
                        "South-East Asia Region" , "Southeast Asia","Southeast Asia, East Asia, and Oceania", "South Asia - World Bank region",
                        "Sub-Saharan Africa - World Bank region" , "Western Europe", "Western Pacific Region" , "Western sub-Saharan Africa", 
                        "World" , "World Bank High Income", "World Bank Low Income","World Bank Lower Middle Income" ,
                        "World Bank Upper Middle Income"))

mean_HIV_AIDS <- HIV_AIDS_nation %>% 
  group_by(Entity)  %>% 
  summarise(`Average Death Rates from HIV/AIDS` =  mean(Deaths_HIV_AIDS, na.rm = TRUE)) %>%
  arrange(-`Average Death Rates from HIV/AIDS`)

mean_HIV_AIDS <- mean_HIV_AIDS %>% 
  mutate_if(is.numeric, round, digits=0)

mean_HIV_AIDS <- head(mean_HIV_AIDS,5)



```


```{r HIVtable,  echo = FALSE, message = FALSE, warning = FALSE}

kable(mean_HIV_AIDS, caption = "Top 5 countries with the highest average death rates from HIV / AIDS",
           booktabs = T) %>%
kable_styling(latex_options = c("striped", "hold_position"))
```

Table \@ref(tab:HIVtable)  above depicts the 5 countries with the highest average death rates from HIV / AIDS from year 1990 to year 2019. South Africa had the highest average death rate, followed by Kenya, Tanzania, India and Nigeria. 

## The death trends for the 5 countries with the highest average death rates from HIV/AIDS
 
Figure \@ref(fig:HIVtrend) depicts the death rates from HIV/AIDS for the 5 countries listed in the table above. Each country appears to follow the global death rate trend of increasing from the 1990s to the 2000s, before decreasing from 2005 to 2020. This is highly evident in South Africa and India. 


```{r HIVtrend, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Top 5 countries with the highest average death rates from HIV/AIDS"  }

HIV_AIDS %>% filter (Entity %in% c("South Africa", "Kenya","Tanzania", "India" , "Nigeria" ))%>%
  ggplot(aes(x= Year, 
             y = Deaths_HIV_AIDS,
             colour = Entity))+
  geom_line(linetype = "dashed", size =1 ) +
  geom_point(size = 2)+
  labs(y = "Number of Deaths from HIV/AIDS",
       x = "Year",
       color = "Country")+
  ggtitle("Deaths rates from HIV/AIDS")+
  scale_y_continuous(labels = scales::comma)

```




\clearpage

## Preventive measures for **HIV / AIDS** 

- *Antiretroviral Therapy (ART)* 

ART treatment involves taking medicines to slow down HIV multiplication rate to prevent HIV from progressing to AIDS, and therefore provide longer lives for patients as discussed in @kemnic2018hiv. Figure \@ref(fig:ARTplot) depicts the rate of HIV/AIDS-related deaths averted due to ART. @sashindran2016antiretroviral and @world2006towards explained that although the treatment was developed in the 1980s, it wasn't widely available in developing countries such as India until the WHO enacted the 3 by 5 initiative between 2003 and 2005. Therefore, figure \@ref(fig:ARTplot) shows an rising trend of HIV/AIDS-related deaths since 2005, which means that the treatment is effective in slowing HIV/AIDS-related deaths. 

This increased the use of ART around the world to treat HIV/AIDS since 2005  can also explain why each country above follows an increasing trend from the mid 2000s. 

```{r ART, echo = FALSE, message = FALSE, warning = FALSE}

ART <- read_csv("Data/hivaids-deaths-and-averted-due-to-art.csv")

ART_nation <- ART %>% filter (Entity %in% c( "South Africa", "Kenya","Tanzania", "India" , "Nigeria" ))%>%
  mutate(Deaths_averted = `Deaths averted due to ART - estimate`) 

``` 



```{r ARTplot, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Effectiveness of Antiretroviral therapy on HIV / AIDS over the years" }

ART_nation %>% 
  
  
  ggplot(aes(x= Year, 
                          y = Deaths_averted, 
                         colour = Entity))+
  geom_line(linetype = "dashed", size =1 ) +
  geom_point( size = 2)+
  labs(y = "HIV/AIDS-related Deaths Averted Due to ART",
       x = "Year")+
  ggtitle("HIV/AIDS-Related Deaths Averted Due to ART")+
  scale_y_continuous(labels = scales::comma)+
  theme_minimal()
  
```

\clearpage

- *Contraception measures*

To further prevent HIV/AIDS, contraception needs to be 'at the centre of the HIV prevention agenda', as discussed in @crankshaw2016placing. @newell1998mechanisms explained that this not only protects people from getting HIV/AIDS but also newborn babies, as HIV/AIDS can be transmitted from mothers to their newborn children.  

- *Education*

Education also needs to be at the forefront of preventive measures as many people in developing countries are unaware of preventative measures for HIV/AIDS, @baxen2004researching.

\section*{}

## Sanna_Meer



\section*{}

## Aphiaut_Imuan 


\clearpage

# Conclusion :


\clearpage