---
title: "Causes of death"
author:
- familyname: Cheung
  othernames: Yan Chui Lucia
  address: Monash Universtidy
  email: yche0164@student.monash.edu
  correspondingauthor: true
  qualifications:  Master of Business Analytics
- familyname: Soh
  othernames: Min Min
  address: Monash Universtidy
  email : mmsoh1@student.monash.edu
  qualifications: Master of Business Analytics
- familyname: Meer
  othernames: Sanna
  address: Monash University
  email: smee0005@student.monash.edu
  correspondingauthor: true
  qualifications:  Master of Business Analytics
- familyname: Imuan
  othernames: Aphiaut
  address: Monash Universtidy
  email: aimu0001@student.monash.edu
  correspondingauthor: true
  qualifications:  Master of Business Analytics
department: ETC5513 Group 4
organization: Monash University
bibliography: References.bib
biblio-style: authoryear-comp
linestretch: 1.5
output:
  bookdown::pdf_document2:
    template: monashreport2.tex
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    includes:
      in_header: preamble.tex
    keep_tex: yes
    number_sections: yes
    citation_package: biblatex
    toc: false
---

```{r , echo = FALSE, message = FALSE, warning = FALSE}
# Libraries
library(tidyverse)
library(readr)
library(kableExtra)
library(bookdown)
library(gridExtra)
library(tinytex)
library(knitr)
```


# Introduction 



# Library used

The library packages used in our research consist of: 

- tidyverse (@tidyverse)
- readr (@readr)
- kableExtra  (@kableExtra)
- Bookdown (@bookdown @bookdown1)
- gridExtra (@gridExtra)
- Tinytex (@tinytex @tinytex1)
- Knitr (@knitr @knitr1)

```{r readdata, , echo = FALSE, message = FALSE, warning = FALSE}
Data <- read_csv("Data/annual-number-of-deaths-by-cause.csv")

```

\clearpage

\section*{}

## YanChuiLucia_Cheung


## MinMin_Soh

# Deaths - HIV/AIDS

Research Questions: 

- How is the death rates of **HIV / AIDS** change in the world ? 
- What is the top 5 countries with the highest death rates from HIV / AIDS ? 
- What are the preventive measures for HIV / AIDS ? 

\clearpage

# Death rates of **HIV / AIDS** in the world  

```{r HIV_AIDS,  echo = FALSE, message = FALSE, warning = FALSE }
HIV_AIDS <- Deaths %>% 
  select( Entity, Code, Year, `Deaths - HIV/AIDS - Sex: Both - Age: All Ages (Number)`) %>%
  rename(Deaths_HIV_AIDS = `Deaths - HIV/AIDS - Sex: Both - Age: All Ages (Number)` )%>%
  filter(!is.na(Deaths_HIV_AIDS))
                
                 
```



```{r Worldtrend, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Number of Deaths from HIV/AIDS in the World" }

HIV_AIDS %>% filter (Entity == "World")%>%
  ggplot(aes(x= Year, 
             y = Deaths_HIV_AIDS))+
  geom_line(linetype = "dashed", color = "blue", size =1 ) +
  geom_point(color = "blue", size = 3)+
  labs(y = "Number of Deaths from HIV/AIDS",
       x = "Year")+
  ggtitle("Deaths rates from HIV/AIDS in the world")+
  scale_y_continuous(labels = scales::comma)+
  theme_minimal()

```

Figure \@ref(fig:Worldtrend)  depicts the global death rate of HIV/AIDS over time. As observed, the death rate followed an increasing trend from 1990 to around 2005, and then decreased from 2005 onwards. 

\clearpage



\clearpage

\section*{}

## Sanna_Meer



\section*{}

## Aphiaut_Imuan 


\clearpage

# Conclusion :


\clearpage