---
title: "Causes of death"
author:
- familyname: Cheung
  othernames: Yan Chui Lucia
  address: Monash Universtidy
  email: yche0164@student.monash.edu
  correspondingauthor: true
  qualifications:  Master of Business Analytics
- familyname: Soh
  othernames: Min Min
  address: Monash Universtidy
  email : mmsoh1@student.monash.edu
  qualifications: Master of Business Analytics
- familyname: Meer
  othernames: Sanna
  address: Monash University
  email: smee0005@student.monash.edu
  correspondingauthor: true
  qualifications:  Master of Business Analytics
- familyname: Imuan
  othernames: Aphiaut
  address: Monash Universtidy
  email: aimu0001@student.monash.edu
  correspondingauthor: true
  qualifications:  Master of Business Analytics
department: ETC5513 Group 4
organization: Monash University
bibliography: References.bib
biblio-style: authoryear-comp
linestretch: 1.5
output:
  bookdown::pdf_document2:
    template: monashreport2.tex
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    includes:
      in_header: preamble.tex
    keep_tex: yes
    number_sections: yes
    citation_package: biblatex
    toc: false
---

```{r , echo = FALSE, message = FALSE, warning = FALSE}
# Libraries
library(tidyverse)
library(readr)
library(kableExtra)
library(bookdown)
library(gridExtra)
library(tinytex)
library(knitr)
```


# Introduction 



# Library used

The library packages used in our research consist of: 

- tidyverse (@tidyverse)
- readr (@readr)
- kableExtra  (@kableExtra)
- Bookdown (@bookdown @bookdown1)
- gridExtra (@gridExtra)
- Tinytex (@tinytex @tinytex1)
- Knitr (@knitr @knitr1)

```{r readdata, , echo = FALSE, message = FALSE, warning = FALSE}
Data <- read_csv("Data/annual-number-of-deaths-by-cause.csv")


```
\clearpage

\section*{Nutrional Deficiencies}

## Sanna_Meer

### Introduction : 
Nutritional deficiency occurs when the body is not getting enough nutrients such as vitamins and minerals. There are a number of conditions that are caused by nutritional deficiency such as Malnutrition or even more severe conditions like Anemia. According to World Health Organisation(WHO), about 2 billion people globally suffer from Nutritional Deficiencies.  
The original dataset is filtered to show details of various countries and the death numbers over these years for Nutritional Deficiencies.  

@khan2010nutritional says that children are the most visible victims of nutritional deficiencies. Poor nutrition contributes to 35% of the 9.2 million child deaths each year globally.

The three research questions that I will look to answer with this data are :  
**How have death from Nutritional Deficiencies changed over time in the world ?**
**What are the top 5 countries with the highest death rates from Nutritional Deficiencies and the death trends for those countries?**
**What are the preventative measures these countries can incorporate to prevent further deaths for Nutritional Deficiencies?**



```{r, echo=FALSE, message=FALSE}
#reading the data
data <- read_csv(here::here("Data/test1.csv"))
data1 <- data %>%
  rename(Nutritional_deficiency_deaths = "Deaths - Nutritional deficiencies - Sex: Both - Age: All Ages (Number)",
         Country = "Entity")

#removing unwanted observations
 data2<- data1 %>% 
   select(Country,Year, Nutritional_deficiency_deaths) %>%
  filter(!Country %in% c("Asia", "Africa", "African Region", "Central Asia", "South-East Asia Region (WHO)", "East Asia & Pacific (WB)", "Region of the Americas (WHO)", "South Asia (WB)",	"African Region (WHO)", "Latin America & Caribbean (WB)", "Eastern Mediterranean Region (WHO)", "Western Pacific Region (WHO)", "OECD Countries", "Sub-Saharan Africa (WB)" ,"Central Europe", "Central Europe, Eastern Europe, and Central Asia",
                        "Central Latin America", "Central sub-Saharan Africa", "Commonwealth High Income", "Commonwealth Low Income", 
                        "Commonwealth Middle Income" , "East Asia","East Asia & Pacific - World Bank region", "Eastern Europe", 
                        "Eastern Mediterranean Region", "Eastern sub-Saharan Africa" ,"Europe & Central Asia - World Bank region", 
                        "European Region", "European Union" , "G20", "Low SDI", "High SDI", "Low-middle SDI",  "High-income", 
                        "High-income Asia Pacific", "High-income North America" , "High-middle SDI","Latin America & Caribbean - World Bank region" ,
                        "Middle SDI","Commonwealth", "African Union", "Southern sub-Saharan Africa", "Region of the Americas",
                        "South-East Asia Region" , "Southeast Asia","Southeast Asia, East Asia, and Oceania", "South Asia - World Bank region",
                        "Sub-Saharan Africa - World Bank region" , "Western Europe", "Western Pacific Region" , "Western sub-Saharan Africa", 
                        "World" , "World Bank High Income", "World Bank Low Income","World Bank Lower Middle Income" ,
                        "World Bank Upper Middle Income"))
 
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}

 Nutr_def_data <- data2 %>%
  select(Country, Year, Nutritional_deficiency_deaths) %>%
  group_by(Country, Year) %>%
  summarise(Total_deaths = sum(Nutritional_deficiency_deaths, na.rm= TRUE))
```


# Worldwide trend


```{r globaltrend, echo = FALSE, message = FALSE, dpi=300, out.width = '70%'}
graph <- data1 %>% filter (Country == "World")

graph1 <-  graph  %>% 
  ggplot(aes(x= Year, 
             y = Nutritional_deficiency_deaths))+
   geom_line(color = "blue") +
  ylab("Total_deaths") + xlab("Year") + ggtitle("The trend of Nutritional Deficiency deaths worldwide from 1990") + scale_y_continuous(labels = scales::comma)

graph1

```

The graph \@ref(fig:globaltrend) shows the worldwide trend of Nutritional Deficiencies from the year 1990. The trend is downward which means that the number of deaths have decreased over these years.


@perez2003nutrition mentions that Health promotion from the early stages in life by fostering healthy eating practices and regular physical activity has had the potential for a major impact on health and well-being during childhood and later stages in life resulting in decreasing deaths. 
This article also mentions that Other features that contribute to gain effectiveness are the provision of adequate time and intensity for the intervention, involvement of families, particularly for younger children, and incorporation of self-assessment and feedback in interventions for older children.  

# Top 5 countries with highest average  death rates


```{r top5, echo = FALSE, message = FALSE}
#top  countries with highest average  death rates

Highest_countries1 <- data2 %>%
         group_by(Country) %>%
          summarise(Average_deaths = mean(Nutritional_deficiency_deaths, na.rm = TRUE)) %>%
  arrange(-Average_deaths)

```


```{r top5countries, echo = FALSE, message = FALSE}
Highest_countries <- Highest_countries1 %>% 
  mutate_if(is.numeric, round, digits=0)

knitr::kable(head(
Highest_countries,5), booktabs = TRUE)

```

The table \@ref(tab:top5countries) tabulates the top 5 countries that have had the maximum deaths. India is at the top with a staggeringly high number followed by Ethiopia, Bangladesh, North Korea and Indonesia.



```{r, echo=FALSE}
deaths_top5 <- data1 %>%
  filter(Country %in% c( "India", "Ethiopia", "Bangladesh", "North Korea", "Indonesia"))
```


# Top 5 countries


```{r topcountries, echo = FALSE, dpi=300, out.width = '70%'}
plot <-  deaths_top5 %>%
  group_by(Country) %>%
  ggplot() +
  geom_line(aes(x = Year, y = Nutritional_deficiency_deaths, color = Country)) +
  ylab("Number of death") + xlab("Year") + ggtitle("Top 5 countries with highest death rates due to Nutritional Deficiencies")

plot
```
\@ref(fig:topcountries) graphically represents these countries. India has an exponentially downward decline over the years. This trend is similar for other countries and the death rate has more or less reached the same level for these countries in the recent past.  

## Reasons :  
 
Although the graph is a positive decline in the trends, there are several reasons as to why these countries have suffered the most :
* Poverty, lack of purchasing power, household food insecurity, and limited general knowledge about appropriate nutritional practices.
* Inadequate maternal and child care, household food insecurity, and poor health services in an unhealthy environment.  

## Preventative Measures :  
* Education programs to encourage healthier and more nutritious food choices
* Food-based approaches that increase the availability or affordability of nutrient-rich foods; and,
* National policies to fortify commonly-eaten foods.

# Conclusion :  

The overall trend of Nutritional Deficiency death is a decreasing one. This means that more countries over the world are embracing the concept and importance of a healthy and balanced diet. @bernert2007nutritional conclude that nutritional deficiencies  require rigorous medical follow-up and multi-disciplinary collaboration between nutritionists, surgeons, dieticians and obstetricians.



\clearpage